---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# georustr

<!-- badges: start -->
<!-- badges: end -->

This repo currently experimental code and a minimal R package to test calling Rust code from R.

The code examples are all documented in this README to keep things simple.
To reproduce these examples (and to begin hacking R/Rust code!) you will need to have installed:

1. [Install the Rust toolchain](https://www.rust-lang.org/tools/install), e.g. with 
```bash
curl --proto '=https' --tlsv1.2 -sSf https://sh.rustup.rs | sh
```
2. [Install R](https://cran.r-project.org/)
3. [Install the R package `sf`](https://rtask.thinkr.fr/installation-of-r-4-0-on-ubuntu-20-04-lts-and-tips-for-spatial-packages/)
4. Install the R package `rextendr`, the development version of which can be installed with the following command:
```{r}
remotes::install_github("extendr/rextendr")
```

After you have installed these things you can clone the repo and open the director in an editor of your choice, e.g. with

```bash 
rstudio georustr/georustr.Rproj # open it in RStudio
# or...
code -r georustr/ # open it in VS Code
code -r georustr/src/rust # open the rust crate in VS Code
```

Rebuild it from the root directory with the following command from the R command line:

```{r}
rextendr::document()
```


Load it with:

```{r}
devtools::load_all()
```

You can download the test data with the following command:

```{r, eval=FALSE}
u = "https://github.com/Robinlovelace/georustr/releases/download/v0.0.0.9000/points.csv"
f = basename(u)
if(!file.exists(f)) {
  download.file(url = u, destfile = f)
}
```

After that the following should work:

```{r}
n = 1e5
points_df = data.frame(x = rnorm(n = n), y = rnorm(n))
system.time({
  points_sf = sf::st_as_sf(points_df, coords = c("x", "y"), crs = 4326)
})
```

We can do the full csv to geojson process for a fair test as follows:

```{r, eval=FALSE}
# run once
readr::write_csv(points_df, "points.csv")
```


```{r}
if(file.exists("points.geojson")) file.remove("points.geojson")
system.time({
  csv_to_json_base_r(file_csv = "points.csv")
})
file.exists("points.geojson")
```

```{r}
if(file.exists("points_rust.geojson")) {
  file.remove("points_rust.geojson")
}
system.time({
  csv_to_geojson_rust()
})
file.exists("points_rust.geojson")
```

The results show that, for this simple test, **Rust is more than 2 times faster than equivalent R/GDAL code**.
Depending on your application, much greater speed-ups should be possible but calling Rust code from R.

You can also run the code from the system command line:

```bash
cd src/rust
cargo test --release
```

Verify the time taken to run as follows:

```bash
time cargo test --release
```

Check the outputs are the same (they are):

```{r, eval=FALSE}
sf::read_sf("points.geojson")
sf::read_sf("points_rust.geojson")
```



